Fail Faster

### 2025-03-11 paper/paper.tex
intro comments

\bcp{The current introduction says, roughly, ``We took a look at
  base.quickcheck, noticed some inefficiencies, applied known
  techniques from staged metaprogramming to eliminate them, and
  observed that speed went up.''  I think we can make much stronger
  claims, but I'm not sure about exactly how strong or what belongs in
  the foreground...
  \begin{itemize}
  \item One of the first serious uses of staged metaprogramming in anger?
  \item A use of staged metaprogramming that is novel / challenging in
  itself, in some way?
  \item An analysis of sources of inefficiency across a range of PBT
  frameworks (and a solution that applies to many of them)?
  \item A usable tool that addresses and overcomes some significant
  implementation challenges?
  \item Careful measurements showing where the sources of inefficiency
  are in existing PBT tools, and which ones matter most?
  \item (Your claim here...)?
  \end{itemize}
}

@@

Our solution is \name, a generator library that preserves
\bq’s functionality while improving performance through
\textit{staged metaprogramming} (staging). Staging is a lightweight,
domain-specific compilation technique that allows us to completely eliminate the
runtime overhead of many common generator abstractions. \jwc{This relies on the fact that the code of generators is known statically at compile time} For instance, in
\name, monadic operations are zero-cost.
\name’s staged eliminates unnecessary allocations,
generates static control structures, and leverages unboxed integer libraries to
achieve performance improvements over \texttt{base\_quickcheck}, making it
well-suited to real-world PBT applications.

@@

In summary, we dramatically improve the bug-finding effectiveness of PBT generators
by optimizing them for speed. In speicific, we make the following contributions:
\begin{enumerate}
    \item An empirical analysis of the sources of inefficiency in PBT
    generators. \jwc{And an argument that PBT generator libraries should be engineered for performance}
    \item A library, \name, which offers efficient \jwc{performant?} generator \proposechange{functions}{combinators} through the
    use of staged metaprogramming; this is the first known application of metaprogramming to PBT.
    \item The insight that generators should be compared by program equality 
    \item An evaluation demonstrating the improved performance of generators constructed
    using our library in a controlled comparison.

@@

\subsection{Staged Metaprogramming}

@@

\jwc{Staged metaprogramming has a long history as a technique for eliminating the runtime overhead of abstractions,
specializing programs at an earlier stage to run more efficiently at a later stage.}

>>>>

\bcp{The current introduction says, roughly, ``We took a look at
  base.quickcheck, noticed some inefficiencies, applied known
  techniques from staged metaprogramming to eliminate them, and
  observed that speed went up.''  I think we can make much stronger
  claims, but I'm not sure about exactly how strong or what belongs in
  the foreground...
  \begin{itemize}
  \item One of the first serious uses of staging in anger?
  \item A use of staging that is novel / challenging in
  itself, in some way?
  \item An analysis of sources of inefficiency across a range of PBT
  frameworks (and a solution that applies to many of them)?
  \item A usable tool that addresses and overcomes some significant
  implementation challenges?
  \item Careful measurements showing where the sources of inefficiency
  are in existing PBT tools, and which ones matter most?
  \item (Your claim here...)?
  \end{itemize}
}

\jwc{Agree: I think something like the 3rd is the best reframing. Something
like: ``We demonstrate that, across languages, monadic PBT generator DSLs can
have a significant performance overhead, and present a cross-language technique
for eliminating the overhead while preserving the idiomatic abstraction.
''}

\jwc{Another angle that I'd argue needs to come out more is that we demonstrate that
engineering PBT libraries with performance in mind has significant impact on testing power.
}

@@

Our solution is \name, a generator library that preserves
\bq’s functionality while improving performance through
\textit{multi-stage programming} (staging). Staging is a lightweight,
domain-specific compilation technique that allows us to completely eliminate the
runtime overhead of many common generator abstractions. \jwc{This relies on the fact that the code of generators is known statically at compile time} For instance, in
\name, monadic operations are zero-cost.
\name’s staged eliminates unnecessary allocations,
generates static control structures, and leverages unboxed integer libraries to
achieve performance improvements over \texttt{base\_quickcheck}, making it
well-suited to real-world PBT applications.

@@

In summary, we dramatically improve the bug-finding effectiveness of PBT generators
by optimizing them for speed. In speicific, we make the following contributions:
\begin{enumerate}
    \item An empirical analysis of the sources of inefficiency in PBT
    generators. \jwc{And an argument that PBT generator libraries should be engineered for performance}
    \item A library, \name, which offers efficient \jwc{performant?} generator \proposechange{functions}{combinators} through the
    use of multi-stage programming; this is the first known application of staging to PBT.
    \item The insight that generators should be compared by program equality 
    \item An evaluation demonstrating the improved performance of generators constructed
    using our library in a controlled comparison.

@@

\subsection{Multi-Stage Programming}

Muli-stage programming (also known more simply as staging) is \jwc{finish}

@@

\jwc{Staging has a long history as a technique for eliminating the runtime overhead of abstractions,
specializing programs at an earlier stage to run more efficiently at a later stage.}
